---
title: Assignment 2
author: Luis Nicolas Luarte Rodriguez
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(ggplot2)
library(knitr)
library(kableExtra)
```


```{r}
# load the dataset
dataSet <- as_tibble(mtcars)
head(dataSet)
```

## Compute the following measures for the variable that you selected for plotting in the previous assignment (I used mtcars dataset and the variable 'mpg')

```{r, echo = FALSE}
d <- mtcars$mpg
descriptiveStats <- data.frame(Statistics = c("Mean", "Median", "Range",
					     "Standard Deviation", "Variance",
					     "Percentile 10", "Percentile 90", 
					     "Quartile 1", "Quartile 3",
					     "Inter-quartile range"),
			      Values = c(mean(d), median(d), rev(range(d)) %>% reduce(`-`), sd(d),
					 var(d), quantile(d, .10), quantile(d, .90),
					 quantile(d, .25), quantile(d, .75), IQR(d))
			      )
descriptiveStats %>%
	kable(., "latex", booktabs = T) %>%
	kable_styling(latex_options = c("striped"), position = "center")
```

## What is the percentage of the data between the mean +/- 1 std?
```{r, echo = FALSE}
lower <- mean(d) - sd(d)
upper <- mean(d) + sd(d)
oneSd <- mean(d >= lower & d <= upper)
twoSd <- mean(d >= lower-sd(d) & d <= upper+sd(d))
threeSd <- mean(d >= lower-(2*sd(d)) & d <= upper+(2*sd(d)))
percentOfData <- data.frame(Criteria = c("1 sd", "2 sd", "3 sd"),
			    Percent = c(oneSd, twoSd, threeSd))
percentOfData %>%
	kable(., "latex", booktabs = T) %>%
	kable_styling(latex_options = c("striped"), position = "center")
```

## Is the distribution of your data close to a normal distribution?
If we compare the values obtained with the ones coming from a normal distribution we get the following differences:

```{r, echo = FALSE}
differences <- data.frame(Normal = c(68.27, 95.45, 99.73),
			    MPG = c(75, 93, 100),
			    Difference = c(abs(68.27 - 75.0),
					   abs(95.45 - 93.0),
					   abs(99.73 - 100.0)))
differences %>%
	kable(., "latex", booktabs = T) %>%
	kable_styling(latex_options = c("striped"), position = "center")
```

The differences are quite small, and this is indication that is close enough to a normal distribution

```{r}
# We can test for normality using the shapiro-wilk test
shapiro.test(d)
# The distribution of the data are not significantly different from normal distribution
```




